The shape of kelp: environmental influences on kelp morphology

### 2018-10-18 chapter_2.Rmd
christo's additional figures added. all figures have been converted to png format. All images can be located in the "figures" folder. PDF output updated.

---
title: "The shape of kelp: environmental influences on kelp morphology"
output:
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    toc: yes
  html_notebook:
    fig_caption: yes
    toc: yes
  pdf_document:
    toc: yes
  bookdown::html_document2:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
fontsize: 12pt
geometry: margin=1in
language: Australian
mainfont: PT Serif
monofont: PT Mono
csl: LaTeX/ecography.csl
papersize: A4
sansfont: PT Sans
tables: yes
bibliography: LaTex/mycollection.bib
---

@@

**insert wave rider images here**

@@

The directional sheltering effect of the Cape Peninsula, against the dominant swell direction, (given in Figure 1) is clear observed in the wave exposures presented in Figure 2. The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits. It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world (Norderhaug *et al.*, 2012; Liberti, Carillo and Sannino, 2013; Sundblad *et al.*, 2014). Kelp and other ecosystems adapt to their typical environmental exposure and thus what is classified as extreme abiotic coastal forcings (temperature, winds, waves and currents) will vary around the world depending on the geographical setting.

@@

*Figure 2: Coastal wave energy exposure around the Cape Peninsula for the total 17-year nearshore wave record, (b) only winter months and (c) only summer months. *

@@

*Figure 3: SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula.  *

>>>>

---
title: "The shape of kelp: environmental influences on kelp morphology"
output:
  pdf_document:
    toc: yes
  html_notebook:
    fig_caption: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    toc: yes
  bookdown::html_document2:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
fontsize: 12pt
geometry: margin=1in
language: Australian
mainfont: PT Serif
monofont: PT Mono
csl: LaTeX/ecography.csl
papersize: A4
sansfont: PT Sans
tables: yes
bibliography: LaTex/mycollection.bib
---

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Wave roses of the Cape Point wave rider buoy situated at 18.29oE, 34.2oS in 70m water depth. (a) total wave record, (b) winter waves and (c) summer waves."}

library(png)
library(grid)
library(gridExtra)

img_twr <-  rasterGrob(as.raster(readPNG("figures/total_wave_rose_2.png")), interpolate = FALSE)
img_wwr <-  rasterGrob(as.raster(readPNG("figures/winter_wave_rose_2.png")), interpolate = FALSE)
img_swr <-  rasterGrob(as.raster(readPNG("figures/summer_wave_rose_2.png")), interpolate = FALSE)

grid.arrange(img_twr, img_wwr,img_swr, ncol = 3)

@@

The directional sheltering effect of the Cape Peninsula, against the dominant swell direction, (given in Figure 1) is clear observed in the wave exposures presented in Figure 2. The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits. It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [Norderhaug2012; @Liberti2013; @Sundblad2014]. Kelp and other ecosystems adapt to their typical environmental exposure and thus what is classified as extreme abiotic coastal forcings (temperature, winds, waves and currents) will vary around the world depending on the geographical setting.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Coastal wave energy exposure around the Cape Peninsula for the (a) total 17-year nearshore wave record, (b) only winter months and (c) only summer months."}

img_we_total <-  rasterGrob(as.raster(readPNG("figures/coastal_we_total.png")), interpolate = FALSE)
img_we_winter <-  rasterGrob(as.raster(readPNG("figures/coastal_we_winter.png")), interpolate = FALSE)
img_we_summer <-  rasterGrob(as.raster(readPNG("figures/coastal_we_summer.png")), interpolate = FALSE)

grid.arrange(img_we_total, img_we_winter,img_we_summer, ncol = 2,nrow = 2)

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="SWAN numerical output for an example typical offshore wave conditions refracting around the Cape Peninsula."}

knitr::include_graphics('figures/regional_waves.png')
